### 2022-01-10 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@461578cc1bc819deaaa27cf8d194d3fe39ed926d 🚀

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily Validating machine learning predictions of
species interactions}
\vskip 5em

@@

\begin{enumerate}
    \item The prediction of species interactions is gaining momentum as
a way to circumvent limitations in data volume. Yet, ecological networks
are challenging to predict because they are typically small and sparse.
Dealing with extreme class imbalance is a challenge for most binary
classifiers, and there are currently no guidelines as to how predictive
models can be trained for this specific problem.%
    \item Using simple mathematical arguments and numerical experiments
in which a variety of classifiers (for supervised learning) are trained
on simulated networks, we develop a series of guidelines related to the
choice of measures to use for model selection, and the degree of
unbiasing to apply to the training dataset.%
    \item Classifier accuracy and the ROC-AUC are not informative
measures for the performance of interaction prediction. PR-AUC is a
fairer assesment of performance. In some cases, even standard measures
can lead to selecting a more biased classifier because the effect of
connectance is strong. The amount of correction to apply to the training
dataset depends on network connectance, on the measure to be optimized,
and only weakly on the classifier.%
    \item These results reveal that training machines to predict
networks is a challenging task, and that in virtually all cases, the
composition of the training set needs to be experimented on before
performing the actual training. We discuss these consequences in the
context of the low volume of data.%
\end{enumerate}

@@

The accuracy paradox is the basis of a number of problems in statistical
education, and lies in the fact that, when the desired class is rare, a
model that gets less and less performant will become more and more
accurate and useful, simply by (i) underpredicting true positive cases
and (ii) over-predicting false negatives. In other words, accuracy,
defined as the proportion of predictions that are correct, is often
useless as a measure of how predictive a model is. This is particularly
true in ecological networks; the desired class (presence of an
interaction between two species) is the one we care most about, and by
far the least commmon. Herein lies the core challenge of predicting
species interactions: the extreme imbalance between classes makes the
training of predictive models difficult. In a recent contribution,
Strydom et al. (2021) highlight that predictive models of interactions
can likely be improved by adding information (in the form of,
\emph{e.g.} traits), but that we do not have robust guidelines as to how
the predictive ability of these models should be evaluated, nor about
how the models should be trained. Here, by relying on simple derivations
and a series of simulations, we formulate a number of such guidelines,
specifically for the case of binary classifiers derived from thresholded
values.

@@

There is an immense amount of measures to evaluate the performance of
classification tasks (Ferri et al., 2009). Here we will focus on five of
them with high relevance for imbalanced learning (He \& Ma, 2013); three
threshold metrics (\(\kappa\), informedness, and the Matthews
Correlation Coefficient), and two ranking metrics (the ROC-AUC and
PR-AUC). The \(\kappa\) measure of agreement (Landis \& Koch, 1977)
establishes the extent to which two observers (here the data and the
prediction) agree, and is measured as

@@

A lot of binary classifiers are built by using a regressor (whose task
is to guess the value of the interaction, amd can therefore return
somethins considered to be a pseudo-probability); in this case, the
optimal value below which predicitions are assumed to be negative
(\emph{i.e.} the interaction does not exist) can be determined by
picking a threshold maximizing some value on the ROC curve or the PR
curve. The area under these curves (ROC-AUC and PR-AUC henceforth) give
ideas on the overall goodness of the classifier. Saito \& Rehmsmeier
(2015) established that the ROC-AUC is biased towards over-estimating
performance for imbalanced data; on the contrary, the PR-AUC is able to
identify classifiers that are less able to detect positive interactions
correctly, with the additional advantage of having a baseline value
equal to prevalence. Therefore, it is important to assess whether these
two measures return different results when applied to ecological network
prediction. The ROC curve is defined by the false positive rate on the
\(x\) axis, and the true positive rate on the \(y\) axis, and the PR
curve is defined by the true positive rate on the \(x\) axis, and the
positive predictive value on the \(y\) axis. By comparison with the
previous paragraph, it is obvious that \(F_1\) has ties to the PR curve
(being close to the expected PR-AUC), and that informedness has ties to
the ROC curve (whereby the threshold maximizing informedness is also the
point of maximal inflection on the ROC curve). One important difference
between ROC and PR is that the later does not prominently account for
the size of the true negative compartments: in short, it is more
sensitive to the correct positive predictions. In a context of strong
imbalance, PR-AUC is therefore a more stringent test of model
performance.

The same approach is used to evaluate \emph{e.g.} species distribution
models (SDMs). Indeed, the training and evaluation of SDMs as binary
classifiers suffers from the same issue of low prevalence. In previous
work, Allouche et al. (2006) suggested that \(\kappa\) was a better test
of model performance than the True Skill Statistic (TSS; which we refer
to as Youden's informedness); these conclusions were later criticized by
Somodi et al. (2017), who emphasized that informedness' relationship to
prevalence depends on assumptions about bias in the model, and therefore
recommend the use of \(\kappa\) as a validation of classification
performance. Although this work offers recommendations about the
comparison of models, it doesn't establishes baselines or good practices
for training on imbalanced ecological data. Within the context of
networks, there are three specific issues that need to be adressed.
First, what values of performance measures are we expecting for a
classifier that has poor performance? This is particularly important as
it can evaluate whether low prevalence can lull us into a false sense of
predictive accuracy. Second, independently of the question of model
evaluation, is low prevalence an issue for \emph{training}, and can we
remedy it? Finally, because the low amount of data on interaction makes
a lot of imbalance correction methods (see \emph{e.g.} Branco et al.,
2015) hard to apply, which indicators can be optimized with the least
amount of positive interaction data?

@@

We establish that due to the low prevalence of interactions, even poor
classifiers applied to food web data will reach a high accuracy; this is
because the measure is dominated by the accidental correct predictions
of negatives. On simulated confusion matrices with ranges of imbalance
that are credible for ecological networks, MCC had the most desirable
behavior, and informedness is a linear measure of classifier skill. By
performing simulations with four models and an ensemble,we show that
informedness and ROC-AUC are consistently high on network data, and that
MCC and PR-AUC are more accurate measures of the effective performance
of the classifier. Finally, by measuring the structure of predicted
networks, we highlight an interesting paradox: the models with the best
performance measures are not the models with the closest reconstructed
network structure. We discuss these results in the context of
establishing guidelines for the prediction of ecological interactions.

@@

In this section, we will assume a network of connectance \(\rho\),
\emph{i.e.} having \(\rho S^2\) interactions (where \(S\) is the species
richness), and \((1-\rho) S^2\) non-interactions. Therefore, the vector
describing the \emph{true} state of the network is a column vector
\(\mathbf{o}^T = [\rho (1-\rho)]\) (we can safely drop the \(S^2\)
terms, as we will work on the confusion matrix, which ends up expressing
\emph{relative} values). We will apply skill and bias to this matrix,
and measure how a selection of performance metrics respond to changes in
these values, in order to assess their suitability for model evaluation.

@@

These two analyses point to the following recommendations: MCC is indeed
more appropriate than \(\kappa\), as although sensitive to bias, it is
sensitive in a consistent way. Informedness is appropriate at
discriminating between different skills, but confounded by bias. \(F_1\)
is increasing with bias, and should not be prioritized to evalue the
performance of the model. The discussion of sensitivity to bias should
come with a domain-specific caveat: although it is likely that
interactions documented in ecological networks are correct, a lot of
non-interactions are simply unobserved; as predictive models are used
for data-inflation (\emph{i.e.} the prediction of new interactions), it
is not necessarilly a bad thing in practice to select models that
predict more interactions than the original dataset, because the
original dataset misses some interactions. Furthermore, the weight of
positive interactions could be adjusted if some information about the
extent of undersampling exists (\emph{e.g.} Branco et al., 2015).

@@

In the following section, we will generate random bipartite networks
(this works without loss of generality on unipartite networks), and
train four binary classifiers (as well as an ensemble model using the
sum of ranged outputs from the component models) on 30\% of the
interaction data. Networks are generated by picking a random
infectiousness trait \(v_i\) for 100 species (from a \(B(6,8)\)
distribution), and a resistance trait \(h_j\) for 100 species (from a
\(B(2,8)\) distribution). There is an interaction between \(i\) and
\(j\) when \(v_i-\xi/2 \le h_j \le v_i+\xi/2\), where \(\xi\) is a
constant regulating the connectance of the network (there is an almost
1:1 relationship between \(\xi\) and connectance), and varies uniformly
in \([0.05, 0.35]\). This model gives fully interval networks that are
close analogues to the bacteria--phage model of Weitz et al. (2005),
with both a modular structure and a non-uniform degree distribution.
This model is easy to learn: when trained with features
\([v_i, h_j, \text{abs}(v_i, h_j)] ^T\) to predict the interactions
between \(i\) and \(j\), all four models presented below were able to
reach almost perfect predictions all the time (data not presented here)
-- this is in part because the rule is fixed for all interactions. In
order to make the problem more difficult to solve, we use \([v_i, h_j]\)
as a feature vector, and therefore the models will have to uncover that
the rule for interaction is \(\text{abs}(v_i, h_j) \le \xi\).

@@

The dataset used for numerical experiments is composed of 64000 such
\((\xi, \nu)\) pairs, on which four machines are trained: a decision
tree regressor, a boosted regression tree, a ridge regressor, and a
random forest regressor. All models were taken from the \texttt{MLJ.jl}
package (Blaom et al., 2020; Blaom \& Vollmer, 2020) in Julia 1.7
(Bezanson et al., 2017). All machines use the default parameterization;
this is an obvious deviation from best practices, as the hyperparameters
of any machine require training before its application on a real
dataset. As we use 64000 such datasets, this would require 256000 unique
instances of tweaking thehyperparameters, which is not realistic.
Therefore, we assume that the default parameterizations are comparable
across networks. All machines return a quantitative prediction, usually
(but not necessarilly) in \([0,1]\), which is proportional (but not
necessarilly linearly) to the probability of an interaction between
\(i\) and \(j\).

@@

Based on the results presented in fig.~\ref{fig:biasmccinf} and
fig.~\ref{fig:biasrocpr}, it seems that informedness and ROC-AUC are not
necessarilly able to discriminate between good and bad classifiers
(although this result may be an artifact for informedness, as it has
been optimized when thresholding). On the other hand, MCC and PR-AUC
show a strong response to training set bias, and may therefore be more
useful at model comparison.

@@

The trained models were then thresholded (again by optimising
informedness), and their predictions transformed back into networks for
analysis; specifically, we measured the connectance, nestedness (REF),
and modularity (REF). This process was repeated 250 times, and the
results are presented in tbl.~\ref{tbl:comparison}. The random forest
model is an interesting instance here: it produces the network that
looks the most like the original dataset, despite having a very low
PR-AUC, suggesting it hits high recall at the cost of low precision.
Although the ensemble was about to reach a very high PR-AUC (and a very
high ROC-AUC), this did not necessarilly translate into more accurate
reconstructions of the structure of the network. This result bears
elaborating. Measures of model performance capture how much of the
interactions and non-interactions are correctly identified. As long as
these predictions are not perfect, some interactions will be predicted
at the ``wrong'' position in the network; these measures cannot describe
the structural effect of these mistakes. On the other hand, measures of
network structure can have the same value with interactions that fall at
drastically different positions; this is in part because a lot of these
measures covary with connectance, and in part because as long as these
values are not 0 or their respective maximum, there is a large number of
network configurations that can have the same value. That ROC-AUC is
consistently larger than PR-AUC may be a case of this measure masking
models that are not, individually, strong predictors (Jeni et al.,
2013).

@@

The results presented here highlight an interesting paradox: although
the Random Forest was ultimately able to get a correct estimate of
network structure tbl.~\ref{tbl:comparison}, it ultimately remains a
poor classifier, as evidenced by its low PR-AUC. This suggests that the
goal of predicting \emph{interactions} and predicting \emph{networks}
may not be solvable in the same way -- of course a perfect classifier of
interactions would make a perfect network prediction; but even the best
scoring predictor of interactions (the ensemble model) had not
necessarilly the best prediction of network structure. The tasks of
predicting networks structure and of predicting interactions within
networks are essentially two different ones. For some applications
(\emph{.e.g.} comparison of network structure across gradients), one may
care more about a robust estimate of the structure, at the cost at
putting some interactions at the wrong place. For other applications
(\emph{e.g.} identifying pairs of interacting species), one may
conversely care more about getting as many pairs right, even though the
mistakes accumulate in the form of a slightly worse estimate of network
structure. How these two approaches can be reconciled is undoubtedly a
task for further research. Despite this apprent tension at the heart of
the predictive exercise, we can use the results presented here to
suggest a number of guidelines.

@@

\textbf{Acknowledgements:} We acknowledge that this study was conducted
on land within the traditional unceded territory of the Saint Lawrence
Iroquoian, Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak
nations. This research was enabled in part by support provided by Calcul
Québec (www.calculquebec.ca) and Compute Canada (www.computecanada.ca)
through the Narval general purpose cluster. TP is supported by a NSERC
Discovery Grant and Discovery Acceleration Supplement, and by a grant
from the Institut de Valorisation des Données (IVADO).

@@

>>>>

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily Guidelines for the validation of machine
learning predictions of species interactions}
\vskip 5em

@@

\begin{enumerate}
    \item The prediction of species interactions is gaining momentum as
a way to circumvent limitations in data volume. Yet, ecological networks
are challenging to predict because they are typically small and sparse.
Dealing with extreme class imbalance is a challenge for most binary
classifiers, and there are currently no guidelines as to how predictive
models can be trained for this specific problem.%
    \item Using simple mathematical arguments and numerical experiments
in which a variety of classifiers (for supervised learning) are trained
on simulated networks, we develop a series of guidelines related to the
choice of measures to use for model selection, and the degree of
unbiasing to apply to the training dataset.%
    \item Neither classifier accuracy nor the ROC-AUC are informative
measures for the performance of interaction prediction. PR-AUC is a
fairer assessment of performance. In some cases, even standard measures
can lead to selecting a more biased classifier because the effect of
connectance is strong. The amount of correction to apply to the training
dataset depends on network connectance, on the measure to be optimized,
and only weakly on the classifier.%
    \item These results reveal that training machines to predict
networks is a challenging task, and that in virtually all cases, the
composition of the training set needs to be experimented on before
performing the actual training. We discuss these consequences in the
context of the low volume of data.%
\end{enumerate}

@@

The accuracy paradox is the basis of a number of problems in statistical
education, and lies in the fact that, when the desired class is rare, a
model that gets less and less performant will become more and more
accurate and useful, simply by (i) underpredicting true positive cases
and (ii) over-predicting false negatives. In other words, accuracy,
defined as the proportion of predictions that are correct, is often
useless as a measure of how predictive a model is. This is particularly
true in ecological networks; the desired class (presence of an
interaction between two species) is the one we care most about, and by
far the least commmon. Herein lies the core challenge of predicting
species interactions: the extreme imbalance between classes makes the
training of predictive models difficult. The connectance of empirical
networks is usually well under 20\%, with larger networks having a lower
connectance (MacDonald et al., 2020). Recent contributions (Becker et
al., 2021; Strydom et al., 2021) highlight that predictive models of
interactions can likely be improved by adding information (in the form
of, \emph{e.g.} traits), but that we do not have robust guidelines as to
how the predictive ability of these models should be evaluated, nor
about how the models should be trained. Here, by relying on simple
derivations and a series of simulations, we formulate a number of such
guidelines, specifically for the case of binary classifiers derived from
thresholded values.

@@

There is an immense diversity of measures to evaluate the performance of
classification tasks (Ferri et al., 2009). Here we will focus on five of
them with high relevance for imbalanced learning (He \& Ma, 2013); three
threshold metrics (\(\kappa\), informedness, and MMC, the Matthews
Correlation Coefficient), and two ranking metrics (the areas under the
Receiving Operator Characteristic and the Precision-Recall curves; resp.
ROC-ACU and PR-AUC). The \(\kappa\) measure of agreement (Landis \&
Koch, 1977) establishes the extent to which two observers (here the data
and the prediction) agree, and is measured as

@@

A lot of binary classifiers are built by using a regressor (whose task
is to guess the value of the interaction, and can therefore return a
value considered to be a pseudo-probability); in this case, the optimal
value below which predictions are assumed to be negative (\emph{i.e.}
the interaction does not exist) can be determined by picking a threshold
maximizing some value on the ROC curve or the PR curve. The area under
these curves (ROC-AUC and PR-AUC henceforth) give ideas on the overall
goodness of the classifier. Saito \& Rehmsmeier (2015) established that
the ROC-AUC is biased towards over-estimating performance for imbalanced
data; on the contrary, the PR-AUC is able to identify classifiers that
are less able to detect positive interactions correctly, with the
additional advantage of having a baseline value equal to prevalence.
Therefore, it is important to assess whether these two measures return
different results when applied to ecological network prediction. The ROC
curve is defined by the false positive rate on the \(x\) axis, and the
true positive rate on the \(y\) axis, and the PR curve is defined by the
true positive rate on the \(x\) axis, and the positive predictive value
on the \(y\) axis. By comparison with the previous paragraph, it is
obvious that \(F_1\) has ties to the PR curve (being close to the
expected PR-AUC), and that informedness has ties to the ROC curve
(whereby the threshold maximizing informedness is also the point of
maximal inflection on the ROC curve). One important difference between
ROC and PR is that the later does not prominently account for the size
of the true negative compartments: in short, it is more sensitive to the
correct positive predictions. In a context of strong imbalance, PR-AUC
is therefore a more stringent test of model performance.

The same approach is used to evaluate \emph{e.g.} species distribution
models (SDMs). Indeed, the training and evaluation of SDMs as binary
classifiers suffers from the same issue of low prevalence; this is not
surprising that the two fields (SDMs and network predictions) would
share methods and their attached conceptual issues, as they suffer from
data limitations, class imbalance, and the conversion of quantitative
prediction into a binary classification. In previous work, Allouche et
al. (2006) suggested that \(\kappa\) was a better test of model
performance than the True Skill Statistic (TSS; which we refer to as
Youden's informedness); these conclusions were later criticized by
Somodi et al. (2017), who emphasized that informedness' relationship to
prevalence depends on assumptions about bias in the model, and therefore
recommend the use of \(\kappa\) as a validation of classification
performance. Although this work offers recommendations about the
comparison of models, it doesn't establishes baselines or good practices
for training on imbalanced ecological data. Within the context of
networks, there are three specific issues that need to be adressed.
First, what values of performance measures are we expecting for a
classifier that has poor performance? This is particularly important as
it can evaluate whether low prevalence can lull us into a false sense of
predictive accuracy. Second, independently of the question of model
evaluation, is low prevalence an issue for \emph{training}, and can we
remedy it? Finally, because the low amount of data on interaction makes
a lot of imbalance correction methods (see \emph{e.g.} Branco et al.,
2015) hard to apply, which indicators can be optimized with the least
amount of positive interaction data?

@@

We establish that due to the low prevalence of interactions, even poor
classifiers applied to food web data will reach a high accuracy; this is
because the measure is dominated by the accidentally correct predictions
of negatives. On simulated confusion matrices with ranges of imbalance
that are credible for ecological networks, MCC had the most desirable
behavior, and informedness is a linear measure of classifier skill. By
performing simulations with four models and an ensemble, we show that
informedness and ROC-AUC are consistently high on network data, and that
MCC and PR-AUC are more accurate measures of the effective performance
of the classifier. Finally, by measuring the structure of predicted
networks, we highlight an interesting paradox: the models with the best
performance measures are not the models with the closest reconstructed
network structure. We discuss these results in the context of
establishing guidelines for the prediction of ecological interactions.

@@

In this section, we will assume a network of connectance \(\rho\),
\emph{i.e.} having \(\rho S^2\) interactions (where \(S\) is the species
richness), and \((1-\rho) S^2\) non-interactions. Therefore, the vector
describing the \emph{true} state of the network (assumed to be an
unweighted, directed network) is a column vector
\(\mathbf{o}^T = [\rho (1-\rho)]\) (we can safely drop the \(S^2\)
terms, as we will work on the confusion matrix, which ends up expressing
\emph{relative} values). We will apply skill and bias to this matrix,
and measure how a selection of performance metrics respond to changes in
these values, in order to assess their suitability for model evaluation.

@@

These two analyses point to the following recommendations: MCC is indeed
more appropriate than \(\kappa\), as although sensitive to bias, it is
sensitive in a consistent way. Informedness is appropriate at
discriminating between different skills, but confounded by bias. \(F_1\)
is increasing with bias, and should not be prioritized to evalue the
performance of the model. The discussion of sensitivity to bias should
come with a domain-specific caveat: although it is likely that
interactions documented in ecological networks are correct, a lot of
non-interactions are simply unobserved; as predictive models are used
for data-inflation (\emph{i.e.} the prediction of new interactions), it
is not necessarily a bad thing in practice to select models that predict
more interactions than the original dataset, because the original
dataset misses some interactions. Furthermore, the weight of positive
interactions could be adjusted if some information about the extent of
undersampling exists (\emph{e.g.} Branco et al., 2015). In a recent
large-scale imputation of interactions in the mammal-virus networks,
Poisot et al. (2021) for example estimated that 93\% of interactions are
yet to be documented.

@@

In the following section, we will generate random bipartite networks
(this works without loss of generality on unipartite networks), and
train four binary classifiers (as well as an ensemble model using the
sum of ranged outputs from the component models) on 30\% of the
interaction data. Networks are generated by picking a random
infectiousness trait \(v_i\) for 100 species (from a \(B(6,8)\)
distribution), and a resistance trait \(h_j\) for 100 species (from a
\(B(2,8)\) distribution). There is an interaction between \(i\) and
\(j\) when \(v_i-\xi/2 \le h_j \le v_i+\xi/2\), where \(\xi\) is a
constant regulating the connectance of the network (there is an almost
1:1 relationship between \(\xi\) and connectance), and varies uniformly
in \([0.05, 0.35]\). This model gives fully interval networks that are
close analogues to the bacteria--phage model of Weitz et al. (2005),
with both a modular structure and a non-uniform degree distribution.
This model is easy to learn: when trained with features
\([v_i, h_j, \text{abs}(v_i, h_j)] ^T\) to predict the interactions
between \(i\) and \(j\), all four models presented below were able to
reach almost perfect predictions all the time (data not presented here)
-- this is in part because the rule is fixed for all interactions. In
order to make the problem more difficult to solve, we use \([v_i, h_j]\)
as a feature vector (\emph{i.e.} the traits on which the models are
trained), and therefore the models will have to uncover that the rule
for interaction is \(\text{abs}(v_i, h_j) \le \xi\).

@@

The dataset used for numerical experiments is composed of 64000 such
\((\xi, \nu)\) pairs, on which four machines are trained: a decision
tree regressor, a boosted regression tree, a ridge regressor, and a
random forest regressor. All models were taken from the \texttt{MLJ.jl}
package (Blaom et al., 2020; Blaom \& Vollmer, 2020) in Julia 1.7
(Bezanson et al., 2017). All machines use the default parameterization;
this is an obvious deviation from best practices, as the hyperparameters
of any machine require training before its application on a real
dataset. As we use 64000 such datasets, this would require 256000 unique
instances of tweaking the hyperparameters, which is not realistic.
Therefore, we assume that the default parameterizations are comparable
across networks. All machines return a quantitative prediction, usually
(but not necessarily) in \([0,1]\), which is proportional (but not
necessarily linearly) to the probability of an interaction between \(i\)
and \(j\).

@@

Based on the results presented in fig.~\ref{fig:biasmccinf} and
fig.~\ref{fig:biasrocpr}, it seems that informedness and ROC-AUC are not
necessarily able to discriminate between good and bad classifiers
(although this result may be an artifact for informedness, as it has
been optimized when thresholding). On the other hand, MCC and PR-AUC
show a strong response to training set bias, and may therefore be more
useful at model comparison.

@@

The trained models were then thresholded (again by optimising
informedness), and their predictions transformed back into networks for
analysis; specifically, we measured the connectance, nestedness (REF),
and modularity (REF). This process was repeated 250 times, and the
results are presented in tbl.~\ref{tbl:comparison}. The random forest
model is an interesting instance here: it produces the network that
looks the most like the original dataset, despite having a very low
PR-AUC, suggesting it hits high recall at the cost of low precision.
Although the ensemble was able to reach a very high PR-AUC (and a very
high ROC-AUC), this did not necessarily translate into more accurate
reconstructions of the structure of the network. This result bears
elaborating. Measures of model performance capture how much of the
interactions and non-interactions are correctly identified. As long as
these predictions are not perfect, some interactions will be predicted
at the ``wrong'' position in the network; these measures cannot describe
the structural effect of these mistakes. On the other hand, measures of
network structure can have the same value with interactions that fall at
drastically different positions; this is in part because a lot of these
measures covary with connectance, and in part because as long as these
values are not 0 or their respective maximum, there is a large number of
network configurations that can have the same value. That ROC-AUC is
consistently larger than PR-AUC may be a case of this measure masking
models that are not, individually, strong predictors (Jeni et al.,
2013).

@@

The results presented here highlight an interesting paradox: although
the Random Forest was ultimately able to get a correct estimate of
network structure tbl.~\ref{tbl:comparison}, it ultimately remains a
poor classifier, as evidenced by its low PR-AUC. This suggests that the
goal of predicting \emph{interactions} and predicting \emph{networks}
may not be solvable in the same way -- of course a perfect classifier of
interactions would make a perfect network prediction; but even the best
scoring predictor of interactions (the ensemble model) had not
necessarily the best prediction of network structure. The tasks of
predicting networks structure and of predicting interactions within
networks are essentially two different ones. For some applications
(\emph{.e.g.} comparison of network structure across gradients), one may
care more about a robust estimate of the structure, at the cost at
putting some interactions at the wrong place. For other applications
(\emph{e.g.} identifying pairs of interacting species), one may
conversely care more about getting as many pairs right, even though the
mistakes accumulate in the form of a slightly worse estimate of network
structure. How these two approaches can be reconciled is undoubtedly a
task for further research. Despite this apparent tension at the heart of
the predictive exercise, we can use the results presented here to
suggest a number of guidelines.

@@

\textbf{Acknowledgements:} We acknowledge that this study was conducted
on land within the traditional unceded territory of the Saint Lawrence
Iroquoian, Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak
nations. This research was enabled in part by support provided by Calcul
Québec (www.calculquebec.ca) and Compute Canada (www.computecanada.ca)
through the Narval general purpose cluster. TP is supported by a NSERC
Discovery Grant and Discovery Acceleration Supplement, by funding to the
Viral Emergence Research Initiative (VERENA) consortium including NSF
BII 2021909, and by a grant from the Institut de Valorisation des
Données (IVADO).

@@

\leavevmode\hypertarget{ref-Poisot2021ImpMam}{}%
Poisot, T., Ouellet, M.-A., Mollentze, N., Farrell, M. J., Becker, D.
J., Albery, G. F., Gibb, R. J., Seifert, S. N., \& Carlson, C. J. (2021,
May 31). \emph{Imputing the mammalian virome with linear filtering and
singular value decomposition}. \url{http://arxiv.org/abs/2105.14973}
